Automatically proving equivalence by type-safe reflection

### 2016-01-12 paper-JFP-DTP/content/2-basicIdeas.tex
paper continued and corrected. Still roughly 2 pages to write

For an expression $e\ :\ Expr\ \Gamma\ x$, we will say that "$e$ denotes (or encodes) the number $x$ in the context $\Gamma$".
When an expression is a variable, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ \Gamma)$.
Also, the $Zero$ expression denotes the natural number $Z$.
Finally, if $e1$ is an expression encoding the number $x$, and $e2$ is an expression encoding the number $y$, then the expression $Plus\ e1\ e2$ denotes the number $(x + y)$.

@@

We want to write the reduction function on a "correct by construction" way, which means that no additional proof should be required after the definition of the function. Thus, $reduce$ will produce the proof that the new Expr freshly produced has the same interpretation as the original Expr, and this will be made easier by the fact that the data type Expr is now indexed over the real --concrete-- number : a term of type $Expr\ \Gamma\ x$ is the encoding of the number $x$.
Thus, we can write the type of $reduce$ like this : \\
$reduce\ :\ Expr\ \Gamma\ x\ \rightarrow\ (x'\ **\ (Expr\ \Gamma\ x',\ x\ =\ x'))$ \\
The function $reduce$ produces a dependent pair : the new concrete number $x'$, and a pair made of an $Expr\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new --concrete-- numbers are equal.
Note that this function can't simply produce an $Expr\ \Gamma\ x$, because the number on which the resulting expression will be indexed is not necessary syntactically equal to the original number since this equality can use the two available properties. Said differently, even if we can prove $x=x'$ (if the function is correctly defined), we do not have $x \equiv x'$.
And in fact, what really interest us in this function is precisely the proof of $x\ =\ x'$.
The reason is that when we try to automatically prove $x=y$, these proofs $x=x'$ and $y=y'$ will be the crucial part for the construction of the desired proof. \\
\\
We have an expression $e1$ encoding $x$, and an expression $e2$ encoding $y$\footnote{These encodings have to be produced by hand by the user for this little tool example, but for the real collection of tactics, we will program an automatic reflection mechanism}.
We will normalize $e1$, and this will give a new number $x'$, a new expression $e1':Expr\ \Gamma\ x'$, and a proof of $x=x'$. We will do the same with $e2$, and we will get a new number $y'$, an expression $e2':Expr\ \Gamma\ y'$, and a proof of $y=y'$. \\
It is now enough to simply compare $e1'$ and $e2'$ using a standard syntactical equality because these two expressions are in normal form :

@@

Now, if the two normalised expressions $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq by default in Idris, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$.
By rewriting the two equalities $x=x'$ and $y=y'$ (that we obtained during the normalisations) in the new equality $x'=y'$, we can get a proof of $x=y$. This is what the function $buildProof$ is doing.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
buildProof : {x : Nat} -> {y : Nat} -> Expr gam x' -> Expr gam y' 
           -> (x = x') -> (y = y') -> Maybe (x = y)
buildProof e e' lp rp with (eqExpr e e')
  buildProof e e lp rp | Just refl = ?MbuildProof
  buildProof e e' lp rp | Nothing = Nothing
\end{verbatim}
\end{center}
\caption{Building the desired proof with the two proofs of equality}
\figrule
\end{figure}

The $Expr\ \Gamma\ x'$ will be used as the normalised left hand size of the equality, which represents the value $x'$. Before the normalisation, the reflected LHS was reflecting the value $x$. The $Expr\ \Gamma\ y'$ is the normalised right hand size, which now represents the value $y'$, but which was encoding $y$ before the normalisation. The function also expects the proofs that $x=x'$ and $y=y'$, and we should be able to pass them because the normalisation function also produces the proof of equality between the previous and the new concrete values.

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
  testEq : Expr gam x -> Expr gam y 
           -> Maybe (x = y)
  testEq l r = 
     let (x' ** (l', p1)) = reduce l in 
     let (y' ** (r', p2)) = reduce r in
        buildProof l' r' p1 p2 
\end{verbatim}
\end{center}
\caption{testEq}
\figrule
\end{figure}

>>>>

For an expression $ex\ :\ Expr\ \Gamma\ x$, we will say that "$ex$ denotes (or encodes) the number $x$ in the context $\Gamma$".
When an expression is a variable, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ \Gamma)$.
Also, the $Zero$ expression denotes the natural number $Z$.
Finally, if $ex$ is an expression encoding the number $x$, and $ey$ is an expression encoding the number $y$, then the expression $Plus\ ex\ ey$ denotes the concrete number $(x + y)$.

@@

We want to write the reduction function on a "correct by construction" way, which means that no additional proof should be required after the definition of the function. Thus, $reduce$ will produce the proof that the new Expr freshly produced has the same interpretation as the original Expr, and this will be made easier by the fact that the data type Expr is now indexed over the real --concrete-- number : a term of type $Expr\ \Gamma\ x$ is the encoding of the number $x$.
Thus, we can write the type of $reduce$ like this : \\
$reduce\ :\ Expr\ \Gamma\ x\ \rightarrow\ (x'\ **\ (Expr\ \Gamma\ x',\ x\ =\ x'))$ \\
The function $reduce$ produces a dependent pair : the new concrete number $x'$, and a pair made of an $Expr\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new --concrete-- numbers are equal.
Note that this function can't simply produce an $Expr\ \Gamma\ x$, because the number on which the resulting expression will be indexed is not necessary syntactically equal to the original number since this equality can use the two available properties. Said differently, even if we can prove $x=x'$ (if the function is correctly defined), we do not have $x \equiv x'$.
And in fact, what really interest us in this function is precisely the proof of $x\ =\ x'$.
The reason is that when we try to automatically prove $x=y$, these proofs $x=x'$ and $y=y'$ will be the crucial part for the construction of the desired proof. \\
\\
We have an expression $ex$ encoding $x$, and an expression $ey$ encoding $y$. We will normalize $ex$, and this will give a new concrete number $x'$, a new expression $ex':Expr\ \Gamma\ x'$, and a proof of $x=x'$. We will do the same with $ey$, and we will get a new concrete number $y'$, an expression $ey':Expr\ \Gamma\ y'$, and a proof of $y=y'$. \\
It is now enough to simply compare $ex'$ and $ey'$ using a standard syntactical equality because these two expressions are now supposed to be in normal form :

@@

Now, if the two normalised expressions $ex'$ and $ey'$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq by default in Idris, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$.
By rewriting the two equalities $x=x'$ and $y=y'$ (that we obtained during the normalisations) in the new equality $x'=y'$, we can get a proof of $x=y$. This is what the function $buildProof$ is doing.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
buildProof : {x : Nat} -> {y : Nat} -> Expr gam x' -> Expr gam y' 
           -> (x = x') -> (y = y') -> Maybe (x = y)
buildProof ex' ey' lp rp with (eqExpr ex' ey')
  buildProof ex' ex' lp rp | Just refl = ?MbuildProof
  buildProof ex' ey' lp rp | Nothing = Nothing
\end{verbatim}
\end{center}
\caption{Building the desired proof with the two proofs of equality}
\figrule
\end{figure}

The $Expr\ \Gamma\ x'$ is the normalised reflected left hand size of the equality, which represents the value $x'$. Before the normalisation, the reflected LHS was reflecting the value $x$. The $Expr\ \Gamma\ y'$ is the normalised reflected right hand size, which now represents the value $y'$, but which was encoding $y$ before the normalisation. The function also expects the proofs that $x=x'$ and $y=y'$, and we should be able to pass them because the normalisation function also produces the proof of equality between the previous and the new concrete values.

@@

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
  testEq : Expr gam x -> Expr gam y 
           -> Maybe (x = y)
  testEq ex ey = 
     let (x' ** (ex', px)) = reduce ex in 
     let (y' ** (ey', py)) = reduce ey in
        buildProof ex' ey' px py 
\end{verbatim}
\end{center}
\caption{testEq}
\figrule
\end{figure}

### 2016-01-12 paper-JFP-DTP/content/3-generalProblem.tex
paper continued and corrected. Still roughly 2 pages to write

All our tactics will require to have a way of testing the equality between elements of the underlying set, that is to say, a way to test equality between constants. For this reason, we define a notion of "set", which only requires this equality. All the algebraic structures above will extend this type class :

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
class Set c where
    set_eq : (x:c) -> (y:c) -> Maybe (x=y)
\end{verbatim}
\end{center}
\caption{Set}
\figrule
\end{figure}

Note that this decidable equality is weak in the sense that it only produces a proof when the two elements are equal, but it doesn't produce a proof of dis-equality when they are different -- instead, it simply produces the value $Nothing$--. That's quite natural, since we want to generate proof of equality, and not to generate counter examples for proving dis-equalities, which is another problem.

Obvisouly, there will be no tactic associated to $Set$, since we have no operations and no properties associated to this structure. Therefore, equalities in a $Set$ are just the syntactical equality, and they can simply be proven with $Refl$.

The first real structure, almost trivial, is the magma. A magma has just a $Plus$ operation which let us computing constants, and has no axioms about this operation.

@@

This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equiped with a $set\_eq$ operation), and if it has a $Plus$ operation.

@@

	\subsection {A bit of notation}
The equality we are trying to prove is $x=y$, where $x$ and $y$ are elements of the type $c$, which  simulates a set with some properties (which make $c$ being a Semi-Group, or a Monoid, or a Group...). The fact that $c$ fulfils the specification of an algebraic structure is expressed as an instance of the corresponding type class, and this instance will be denoted as $p$.
The reflected term for $x$ will be denoted $e1$, and this term will have the type $ExprG\ p\ \Gamma\ x$. The term $e1$ is the encoding of $x$ and its type is precisely indexed over the real value $x$, as the reflected terms embed the concrete values.
We've got similar thing for $y$, which is encoded by $e2$, and its type is indexed over the real value $y$.
Running the normalisation procedure on $e1$ will produce the normal form $e1'$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $p1$ of $x=x'$ while running the normalisation procedure on $e2$ will produce the normal form $e2'$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $p2$ of $y=y'$.

@@

Each algebraic structure will have a function for reducing the reflected terms into their normal forms. The fact that all of these algebraic structures admit a canonical represent for any element is in fact a very nice property that we are using in order to decide equalities. Without this property, it would become a lot more complicated to decide the equality without bruteforcing a serie of rewriting, that might even not terminate.
These functions will compute this canonical represent, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the real value indexing the original term and the real value indexing the produced term.

@@

The first thing this function has to do is to compute the normal form of the two expressions $e1$ and $e2$ respectively encoding $x$ and $y$.
Then, the only thing remaining to do will be to compare syntactically $e1'$ and $e2'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x=y$ by using the two proofs $p1$ and $p2$.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
groupDecideEq p e1 e2 =
  let (x' ** (e1', p1)) = groupReduce p e1 in
  let (y' ** (e2', p2)) = groupReduce p e2 in
	     buildProofGroup p e1' e2' p1 p2
\end{verbatim}
\end{center}
\caption{Decides if two elements of a group are equal}
\figrule
\end{figure}

The syntactical test of equality between $e1'$ and $e2'$ and the composition of the two proofs is done in the auxiliary function $buildProofGroup$, similarly to what we've done on the smaller example in the previous section.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
buildProofGroup : (p:Group c) -> {gam:Vect n c} 
  -> {x : c} -> {y : c} -> {x':c} -> {y':c} 
  -> (ExprG p gam x') -> (ExprG p gam y') 
  -> (x = x') -> (y = y') 
  -> (Maybe (x = y))
buildProofGroup p e1 e2 p1 p2 with (exprG_eq p _ e1 e2)
    buildProofGroup p e1 e1 p1 p2 | Just refl = ?MbuildProofGroup
    buildProofGroup p e1 e2 p1 p2 | Nothing = Nothing
\end{verbatim}
\end{center}
\caption{Composes the two proofs if the normal forms are the same}
\figrule
\end{figure}

>>>>

In fact, we don't only want to be able to prove equalities, but more generally equivalence. We want to allow the user the possibility to define his own equivalence notion that interests him, assuming he's able to provide the properties that belongs to the algebraic structure that he wants to use. This equivalence relation on $c$ has the following profile $(\sim=) : (x:c) \rightarrow (y:c) \rightarrow Type$\footnote{This Type would be a Prop in system like Coq that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.

All of our tactics will require to have a way of testing this equivalence between elements of the underlying set, that is to say, a way to test equivalence between constants. For this reason, we define a notion of "set", which only requires the definition of the equivalence relation --accompanied with the proofs that this is indeed an equivalence relation-- and this equivalence test $set\_eq$. All the algebraic structures will later extend this type class :

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
class Set c where
    (~=) : (x:c) -> (y:c) -> Type

    refl : (x:c) -> x~=x
    sym : {x:c} -> {y:c} -> (x~=y) -> (y~=x)
    trans : {x:c} -> {y:c} -> {z:c} -> (x~=y) -> (y~=z) -> (x~=z)

    set_eq : (x:c) -> (y:c) -> Maybe (x=y)
\end{verbatim}
\end{center}
\caption{Set}
\figrule
\end{figure}
If the user wants to prove propositional equalities, he will simply instantiate $(\sim=)$ with the built-in $(=)$ during the definition of the instance of $Set$.
Note that $(\sim=)$ is only weakly decidable in the sense that $set\_eq$ only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value $Nothing$--. That's quite natural, since we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.

Obviously, there will be no tactic associated to $Set$, since we have no operations and no properties associated to this structure. Therefore, equivalence in a $Set$ are just the "syntactical equivalence", and they can simply be proven with $refl$.

The first real structure, almost trivial, is the magma. A magma has just a $Plus$ operation which let us computing constants, and has no properties about this operation.

@@

This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equiped with the equivalence relation $\sim=$ and the test $set\_eq$), and if it has a $Plus$ operation.

@@

	\subsection {A bit of notation}
The equality we are trying to prove is $x=y$, where $x$ and $y$ are elements of the type $c$, which  simulates a set with some properties (which make $c$ being a Semi-Group, or a Monoid, or a Group...). The fact that $c$ fulfils the specification of an algebraic structure is expressed as an instance of the corresponding type class, and this instance will be denoted as $p$.
The reflected term for $x$ will be denoted $ex$, and this term will have the type $ExprG\ p\ \Gamma\ x$. The term $ex$ is the encoding of $x$ and its type is precisely indexed over the real value $x$, as the reflected terms embed the concrete values.
We've got similar thing for $y$, which is encoded by $ey$, and its type is indexed over the real value $y$.
Running the normalisation procedure on $ex$ will produce the normal form $ex'$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $px$ of $x=x'$ while running the normalisation procedure on $ey$ will produce the normal form $ey'$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $py$ of $y=y'$.

@@

Each algebraic structure will have a function for reducing the reflected terms into their normal forms. The fact that all of these algebraic structures admit a canonical represent for any element is in fact a very nice property that we are using in order to decide equalities. Without this property, it would become a lot more complicated to decide the equality without brute-forcing a serie of rewriting, that might even not terminate.
These functions will compute this canonical represent, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the real value indexing the original term and the real value indexing the produced term.

@@

The first thing this function has to do is to compute the normal form of the two expressions $ex$ and $ey$ respectively encoding $x$ and $y$.
Then, the only thing remaining to do will be to compare syntactically $ex'$ and $ey'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x=y$ by using the two proofs $px$ and $py$.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
groupDecideEq p ex ey =
  let (x' ** (ex', px)) = groupReduce p ex in
  let (y' ** (ey', py)) = groupReduce p ey in
	     buildProofGroup p ex' ey' px py
\end{verbatim}
\end{center}
\caption{Decides if two elements of a group are equal}
\figrule
\end{figure}

The syntactical test of equality between $ex'$ and $ey'$ and the composition of the two proofs is done in the auxiliary function $buildProofGroup$, similarly to what we've done on the smaller example in the previous section.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
buildProofGroup : (p:Group c) -> {gam:Vect n c} 
  -> {x : c} -> {y : c} -> {x':c} -> {y':c} 
  -> (ExprG p gam x') -> (ExprG p gam y') 
  -> (x = x') -> (y = y') 
  -> (Maybe (x = y))
buildProofGroup p ex ey px py with (exprG_eq p _ ex ey)
    buildProofGroup p ex ex px py | Just refl = ?MbuildProofGroup
    buildProofGroup p ex ey px py | Nothing = Nothing
\end{verbatim}
\end{center}
\caption{Composes the two proofs if the normal forms are the same}
\figrule
\end{figure}

### 2016-01-12 paper-JFP-DTP/content/4-detailsNormalisation.tex
paper continued and corrected. Still roughly 2 pages to write

We now describe briefly how the normalisation functions work for some of the structures, and we also explain how we've been able to reuse each prover for the more specialised underneath structures.

	\subsection {Normalization of terms in a Semi-Group}
In a semigroup, we only have to deal with the property of associativity. As it was the case with the toy example in the previous section, we will have to rearrange the parenthesis on a systematical way, either left associative, or right associative. But this is not the only thing that we have to do. If $x$, $y$ and $z$ denote variables, we don't only want to transform $x+((y+4)+(5+z))$ into $(((x+y)+4)+5)+z$ if we've chosen the complete left associative form. We want first to do the addition of the constants $4$ and $5$ together (and we have the right to do so because we can rearrange the parenthesis thanks to the associativity, when constants are, like in this case, near to each other). Three possible patterns need this treatment : $(x+c1)+(c2+y)$ , $(x + c1) + c2$  and $c1 + (c2 + x)$ when $c1$ and $c2$ denote constants. The simplification between the constants $c1$ and $c2$ is denoted here as $[c1+c2]$, and this computation is done by the normalisation function of the underneath $Magma$ level, called $magmaReduce$.
The first pattern is rewritten into $(x + [c1+c2]) + y$, the second into $x + [c1+c2]$ and the third one into $[c1 + c2] + x$.
We show the code for the first pattern.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
assoc : (p:SemiGroup c) -> (gam:Vect n c) -> {c1:c} 
        -> (ExprSG p gam c1) -> (c2 ** (ExprSG p gam c2, c1=c2))
assoc p gam (PlusSG 
             (PlusSG e1 (ConstSG _ _ c1)) 
             (PlusSG (ConstSG _ _ c2) e2)) =
    let (r_ih1 ** (e_ih1, p_ih1)) = (assoc p gam e1) in
    let (r_ih2 ** (e_ih2, p_ih2)) = (assoc p gam e2) in
    let (r_3 ** (e_3, p_3)) = magmaReduce (semiGroup_to_magma 
                             (PlusSG (ConstSG _ _ c1) (ConstSG _ _ c2))) in
    let e_3' = magma_to_semiGroup p e_3 in
        (_ ** ((PlusSG (PlusSG e_ih1 e_3') e_ih2), ?Massoc1))
[...]
\end{verbatim}
\end{center}
\caption{Computing with associativity in a Semi-Group, first pattern}
\figrule
\end{figure}

Note that we use function of conversions between the different levels : $semiGroup\_to\_magma$ and $magma\_to\_semiGroup$.

After simplifying the constants computable thanks to associativity, the normalization for SemiGroup also has to rearrange the brackets on a systematic way. The fully right or fully left associative forms can both be chosen for this purpose, and the implementation is very similar to the function $expr_l$ previously described on section 2.3.

>>>>

We now describe briefly how the normalisation functions work for some of the different algebraic structures, and we also explain how we've been able to reuse each prover for the more specialised underneath structures. By doing so, we show that the construction of the desired proof is always done bit by bit, because every rewriting of the reflected term is immediately justified by a little accompanying proof which shows the equivalence of the previous and new index.
\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
littleRewriting : (p:algebraic structure on c) -> (gam:Vect n c) -> {c1:c} 
        -> (Expr p gam c1) -> (c2 ** (Expr p gam c2, c1=c2))
simpleTreatment p gam e = (new expression, justification)
\end{verbatim}
\end{center}
\caption{General pattern of a function doing a part of the normalisation}
\figrule
\end{figure}

	\subsection {Normalization of terms in a Semi-Group}
In a semigroup, we only have to deal with the property of associativity. As it was the case with the toy example in the previous section, we will have to rearrange the parenthesis on a systematic way, either left associative, or right associative. But this is not the only thing that we have to do. If $x$, $y$ and $z$ denote variables, we don't only want to transform $x+((y+4)+(5+z))$ into $(((x+y)+4)+5)+z$ --we assume we've chosen the complete left associative form--. We also want the constants $4$ and $5$ to be simplified together, because they are close to each other (and we have the right to do so because we can rearrange the parenthesis thanks to the associativity), and the result should be $((x+y)+9)+z$ on this example. Three possible patterns need this treatment : $(x+const1)+(const2+y)$, $(x + const1) + const2$  and $const1 + (const2 + x)$ where $const1$ and $const2$ denote constants. The simplification between the constants $const1$ and $const2$ is denoted here as $[const1+const2]$, and this computation can be done at the underneath $Magma$ level, with its normalisation function called $magmaReduce$.
The first pattern is rewritten into $(x + [const1+const2]) + y$, the second into $x + [const1+const2]$ and the third one into $[const1 + const2] + x$.
We show the code for the first pattern.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
assoc : (p:SemiGroup c) -> (gam:Vect n c) -> {c1:c} 
        -> (ExprSG p gam c1) -> (c2 ** (ExprSG p gam c2, c1=c2))
assoc p gam (PlusSG 
             (PlusSG ex (ConstSG _ _ const1)) 
             (PlusSG (ConstSG _ _ const2) ey)) =
    let (r_ihx ** (e_ihx, p_ihx)) = (assoc p gam ex) in
    let (r_ihy ** (e_ihy, p_ihy)) = (assoc p gam ey) in
    let (r_3 ** (e_3, p_3)) = magmaReduce (semiGroup_to_magma 
                             (PlusSG (ConstSG _ _ const1) (ConstSG _ _ const2))) in
    let e_3' = magma_to_semiGroup p e_3 in
        (_ ** ((PlusSG (PlusSG e_ihx e_3') e_ihy), ?Massoc1))
[...]
\end{verbatim}
\end{center}
\caption{Computing with associativity in a Semi-Group, first pattern}
\figrule
\end{figure}
The treatment is applied recursively on the sub-expressions $ex$ and $ey$, which produces the new expressions $e\_ihx$ and $e\_ihy$ together with the new concrete values they represent ($r\_ihx$ and $r\_ihy$) and the proofs ($p\_ihx$ and $p\_ihy$) justifying these transformation on the concrete values. Note that the simplification $[const1+const2]$ is done by calling the magma prover, which gives a new expression $e3$, even thought it could perfectly be done directly here. We wanted to show how a prover can be reused for building another prover on an easy example. The functions of conversion between the different levels : $semiGroup\_to\_magma$ and $magma\_to\_semiGroup$ are helping for this task.
\\
A meta-variable Massoc1 has been introduced. This metavariable corresponds to a proof obligation for the following property :
$Massoc1 : Plus\ (Plus\ x\ const1)\ (Plus\ const2\ y)\ ~=\ Plus\ (Plus\ r\_ihx\ r_3)\ r\_ihy$ in a context where there is, amongst other things, $p\_ihx : x\ ~=\ r\_ihx$, $p\_ihy : y\ ~=\ r\_ihy$ and $p_3 : Plus\ const1\ const2 ~= r\_3$.
This lemma says that the treatment done by this function $assoc$ on this case was sound because the previous and the new concrete values are equal. It can easily be proven with a few rewritings and the use of the associativity, available from the semi-group structure. \\
\\
After this simplification of constants computable thanks to associativity, the normalization for SemiGroup also has to rearrange the brackets on a systematic way. The fully right or fully left associative forms can both be chosen for this purpose, and the implementation is very similar to the function $expr_l$ previously described for the smaller example on section 2.3.
